% !TEX root = sycl-1.2.1.tex
%*******************************************************************************
% SYCL support of non-core OpenCL features
%*******************************************************************************
\chapter{SYCL Support of Non-Core OpenCL Features}

In addition to the OpenCL core features, SYCL also provides support for OpenCL
extensions which provide features in OpenCL via khr extensions.

Some extensions are natively supported within the SYCL interface, however some
can only be used via the OpenCL interoperability interface. The SYCL interface
required for native extensions must be available. However if the respective
extension is not supported by the executing SYCL \codeinline{device}, the
\gls{sycl-runtime} must throw a \codeinline{feature_not_supported} exception.

A SYCL \codeinline{platform} or SYCL \codeinline{device} can be queried for
extension support by calling the \codeinline{has_extension} member function.

The level of support for the OpenCL extensions are listed in
Table~\ref{table.extensionsupport}.

%-------------------------------------------------------------------------------
\begin{longtable}[!h]
 {|>{\raggedright\arraybackslash}p{2.0 in} | p{1.5 in} | p{1.5 in} |}
  \hline
  \cellcolor{lightgray}\textbf{Extension}
  & \cellcolor{lightgray}\textbf{SYCL interface}
  & \cellcolor{lightgray}\textbf{OpenCL interoperability interface} \\\hline
  \endhead
\addInfoFootNotes{SYCL support for OpenCL 1.2 extensions}
{table.extensionsupport}
  \addInfoRow
    {cl_khr_int64_base_atomics} {Yes} {Yes}
  \addInfoRow
    {cl_khr_int64_extended_atomics} {Yes} {Yes}
  \addInfoRow
    {cl_khr_fp16} {Yes} {Yes}
  \addInfoRow
    {cl_khr_3d_image_writes} {Yes} {Yes}
  \addInfoRow
    {cl_khr_gl_sharing} {No} {Yes}
  \addInfoRow
    {cl_apple_gl_sharing} {No} {Yes}
  \addInfoRow
    {cl_khr_d3d10_sharing} {No} {Yes}
  \addInfoRow
    {cl_khr_d3d11_sharing} {No} {Yes}
  \addInfoRow
    {cl_khr_dx9_media_sharing} {No} {Yes}
\end{longtable}
%-------------------------------------------------------------------------------

%*******************************************************************************
% Half Precision Floating-Point
%*******************************************************************************
\section{Half Precision Floating-Point}
\label{sec:extension-fp16}

The half scalar data type: \codeinline{half} and the half vector data types: \codeinline{half1}, \codeinline{half2}, \codeinline{half3}, \codeinline{half4}, \codeinline{half8} and \codeinline{half16} must be available at compile-time. However if any of the above types are used in a SYCL kernel function, executing on a device which does not support the extension \codeinline{khr_fp16}, the \gls{sycl-runtime} must throw a \codeinline{feature_not_supported} exception.

The conversion rules for half precision types follow the same rules as in the OpenCL 1.2 extensions specification \cite[par. 9.5.1]{openclext12}.

The math functions for half precision types follow the same rules as in the OpenCL 1.2 extensions specification \cite[par. 9.5.2, 9.5.3, 9.5.4, 9.5.5]{openclext12}. The allowed error in ULP(Unit in the Last Place) is less than  8192, corresponding to Table 6.9 of the OpenCL 1.2 specification \cite{opencl12}.

%*******************************************************************************
% 64 bit atomics
%*******************************************************************************
\section{64 Bit Atomics}
\label{sec:extension.64bit-atomics}

The SYCL \codeinline{atomic} class can support 64 bit data types if the 64 bit
extensions \codeinline{cl_khr_int64_base_atomics} and \codeinline{
cl_khr_int64_extended_atomics} are supported by the SYCL \codeinline{device}
which is executing the SYCL kernel function.

The extension \codeinline{cl_khr_int64_base_atomics} adds support for the
\codeinline{load}, \codeinline{store}, \codeinline{fetch_add},
\codeinline{fetch_sub}, \codeinline{exchange} and
\codeinline{compare_exchange_strong} member functions with 64 bit data types.

The extension \codeinline{cl_khr_int64_extended_atomics} adds support for the
\codeinline{fetch_min}, \codeinline{fetch_max}, \codeinline{fetch_and},
\codeinline{fetch_or} and \codeinline{fetch_xor} member functions with 64 bit
data types.

%*******************************************************************************
% Writting to 3D image memory objects
%*******************************************************************************
\section{Writing to 3D image memory objects}

The \codeinline{accessor} class for target
\codeinline{access::target::image} in SYCL support member functions for writing
3D image memory objects, but this functionality is \keyword{only
allowed} on a device if the extension \tf{cl_khr_3d_image_writes} is
supported on that \keyword{device}.

\section{Interoperability with OpenGL}

Interoperability between SYCL and OpenGL is not directly provided by the SYCL interface,
however can be achieved via the SYCL OpenCL interoperability interface.

\clearpage

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "sycl-1.2.1"
%%% TeX-auto-untabify: t
%%% TeX-PDF-mode: t
%%% ispell-local-dictionary: "american"
%%% End:
